### 2019-07-17 manuscript/discussion.tex
Fixed all todos and revised discussion issues

Throughout our numerous model comparisons, we find that inclusion of hidden traits provides a considerably better fit (\cref{supptable:testaddhidden}). 
This is consistent with the expectation that many processes, beyond those associated with the focal traits, can affect inference of speciation and extinction. %removed "one or two" because it's true for N traits.
%A natural follow-up question is, what is this hidden trait?
It is, however, unclear exactly which processes are captured by the hidden traits.
For example, our results show that, to a varying extent, breeding system functions as a hidden trait in the ploidy-centered analysis, and vice versa (\cref{suppfigure:lumpedDP,suppfigure:lumpedIC}).
But the strong statistical support for processes not well explained by ploidy and breeding system (\cref{supptable:testaddhidden}) tempts one to interpret the remaining variation as the effect of other measurable traits.
For example, our data appear to show a rapidly-diversifying Australasian clade of mostly SC species within \emph{Solanum}, which suggests that geography may play a role. % Boris, please check!  Australian?  fixme
Nevertheless, it is also possible that the addition of hidden traits instead explains variation stemming from any of a number of unrelated processes or methodological artifacts, such as variation in rates of molecular evolution or misspecified models of sequence evolution. %Rosana and Emma check me.
%     but don't be circular, test in different clade
% alpha is always low: new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model
%B: this transition from the meaning of hidden states to best practices of lumping is a bit hard.
%Here Rosana adds the value of adding traits with the lumped models
More generally, when analyzing the value of additional traits to a complex model, it may be best to opt for the use of lumped models (\cref{figure:lumped,table:lumped}), and verify that adding a trait is statistically preferred.
In the absence of additional information, the hidden states can be viewed as a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.
In our analyses, the transition rates between and within hidden traits did not affect the direction of the net diversification difference. 
We find that it is important to allow for the hidden state transition rates to be asymmetrical, because those models tend to provide a better fit than models that fix the hidden state transition rates to be equal (\cref{suppfigure:asymmetric,supptable:asymmetry}).

@@

Macroevolutionary transition rates represent a combination of time spent waiting for individuals with a new character state, and for that new state to become widespread within the species.
For our traits, this consists of mutations that break SI or generate polyploid individuals, and selective pressures that cause fixation (or loss) of these mutants.
Estimates of mutation rates are highly uncertain, but the chance of breakdown of SI within diploids is perhaps $10^{-5}$ per pollen grain; this includes breakdown by autopolyploidization \citep{lewis1979}, which is by itself estimated to occur approximately within the same order of magnitude \citep{ramsey_1998}. 
% Seemingly, then, the simple genic mutation rate that leads to loss of SI is at best equal to tetraploidization mutation rate, and possibly far lower.
In contrast, we infer a macroevolutionary transition rate from $ID$ to $CD$ that is three-fold greater than the rate from $ID$ to $CP$, indicating that selection restricts the fixation of new polyploids more than of new SC mutants  (\cref{figure:pathways}, \citealt{robertson_2011}). % todo: check if new rates are still 3x %FIXME

@@

Other lines of evidence about the prevalence of diploidization within Solanaceae or its ancestors are mixed or even conflicting.
% Part I: evidence for WGD in/near family
On the one hand, polyploidy may have occurred prior to the origin of Solanaceae, rendering all extant `diploids' secondarily derived.
\Citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \emph{Solanum lycopersicum}, may have experienced one or more whole genome duplications.
A subsequent analysis of synteny between grape and \emph{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \emph{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}. 
The age of the peak of paralog Ks distances is approximately $71 \pm 19$ My \citep{tomato2012}. 
If this is the case, then all of the genomes may have been subsequently diploidized, yielding the widespread and frequent chromosome numbers in this and closely related families, n=11--12, all presently considered to be diploid \citep{robertson_2011}.

% Part II: evidence against WGD in/near family
On the other hand, there is little evidence for the occurrence of diploidization within Solanaceae itself.
First, the Ks-inferred duplication \citep{tomato2012} likely pre-dates the origin of the family (49 My, HPD 46--53 My; \citealt{sarkinen_2013}). 
Thus, the species in our analyses could be ancestrally polyploid, but would have diploidized before the root of the family.
Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species. % including all of those in the relatively cytogenetically conserved x=12 group, which includes tomato, potato, eggplant, pepper, and tobacco.
Specifically, chromosome numbers and genome comparisons within the family (esp.\ the x=12 clade containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ by as few as six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations.
Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
%
% Part IIb: evidence from SI [might need more explaining?]
Furthermore, whole genome duplication in a eudicot lineage ancestral to Solanaceae would clash with the evidence that the homologous mechanism of SI, which has been present continually in many families \citep{igic_2006}, breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002, mcclure_2009}.
Most problematically in this context, it is unclear how to explain the maintenance of trans-generic polymorphism at the orthologous S-loci in Solanaceae and other families if SI was previously broken down by polyploidization.
%
Regardless of whether genome polyploidization, followed by widespread diploidization, is a dominant mechanism of genome evolution in Solanaceae, it is clear that more work is needed for a complete understanding of the joint evolution of ploidy and breeding systems.

>>>>

Throughout our numerous model comparisons, we find that inclusion of hidden states provides a considerably better fit (\cref{supptable:testaddhidden}). 
This is consistent with the expectation that many processes, beyond those associated with the focal traits, can affect inference of speciation and extinction. %removed "one or two" because it's true for N traits.
%A natural follow-up question is, what is this hidden trait?
It is, however, unclear exactly which processes are captured by the hidden states.
For example, our results show that, to a varying extent, breeding system functions as a hidden state in the ploidy-centered analysis, and vice versa (\cref{suppfigure:lumpedDP,suppfigure:lumpedIC}).
But the strong statistical support for processes not well explained by ploidy and breeding system (\cref{supptable:testaddhidden}) tempts one to interpret the remaining variation as the effect of other measurable traits.
For example, our data appear to show a rapidly-diversifying Australasian clade of mostly SC species within \emph{Solanum}, which suggests that geography may play a role. % Boris, please check!  Australian?  fixme
Nevertheless, it is also possible that the addition of hidden states instead explains variation stemming from any of a number of unrelated processes or methodological artifacts, as previously discussed by \citet{beaulieu_2016}.
In the absence of additional information, the hidden states can be viewed as a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.

%such as variation in rates of molecular evolution or misspecified models of sequence evolution. %Rosana and Emma check me.
%R- let's not get into the weeds of what is a hidden state (I remove the use of trait  for state because a state is what the model receives) . Jeremy does a good job discussing this in his paper so no need.

%     but don't be circular, test in different clade
% alpha is always low: new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model - it is low because not enough transitions and mostly a different process towards the tips-who knows what is happening there?
%B: this transition from the meaning of hidden states to best practices of lumping is a bit hard.- Fixed below
%Here Rosana adds the value of adding traits with the lumped models
As more trait information becomes available for macroevolutionary studies, it is important not only recognizing the role of hidden states as a part of a general modeling approach, but also questioning whether adding more traits to diversification studies is relevant.
More generally, when analyzing the value of adding additional traits to an already complex model with or without hidden states, we find that using lumped models  is a key tool to verify that adding a trait improves our understanding of the diversification patterns (\cref{figure:lumped,table:lumped}).

%R-removed this, it is really not important to discuss
%In our analyses, we find that the transition rates between and within hidden traits did not affect the direction of the net diversification differences. 
%We find that it is important to allow for the hidden state transition rates to be asymmetrical, because those models tend to provide a better fit than models that fix the hidden state transition rates to be equal (\cref{suppfigure:asymmetric,supptable:asymmetry}).

@@

Macroevolutionary transition rates represent a combination of time spent waiting for individuals with a new character state, and for that new state to become widespread within the species.
For our traits, this consists of mutations that break SI or generate polyploid individuals, and selective pressures that cause fixation (or loss) of these mutants.
Estimates of mutation rates are highly uncertain, but the chance of breakdown of SI within diploids is perhaps $10^{-5}$ per pollen grain; this includes breakdown by autopolyploidization \citep{lewis1979}, which is by itself estimated to occur approximately within the same order of magnitude \citep{ramsey_1998}. 
% Seemingly, then, the simple genic mutation rate that leads to loss of SI is at best equal to tetraploidization mutation rate, and possibly far lower.
In contrast, we infer a macroevolutionary transition rate from $ID$ to $CP$ that is 2. 5 times greater than the rate from $ID$ to $CD$, indicating that selection restricts the fixation of new polyploids more than of new SC mutants  (\cref{figure:pathways}, \citealt{robertson_2011}). 
% todo: check if new rates are still 3x 
%R- fixed, it is two and a half more and it was written upside down ID to CP 0.194 ID to CD 0.073 which means more evolution to CP than to CD.

@@

Other lines of evidence about the prevalence of diploidization within Solanaceae or its ancestors are mixed or even conflicting.
% Part I: evidence for WGD in/near family
On the one hand, polyploidy may have occurred prior to the origin of Solanaceae, rendering all extant `diploids' secondarily derived.
\Citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \emph{Solanum lycopersicum}, may have experienced one or more whole genome duplications.
A subsequent analysis of synteny between grape and \emph{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \emph{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication ( $71 \pm 19$ My, \citet{tomato2012})  likely pre-dating the origin of the family ($49 \pm 3$ My; \citealt{sarkinen_2013}).

%The age of the peak of paralog Ks distances is approximately $71 \pm 19$ My \citep{tomato2012}. 
% R- not entering into the weeds of Ks plots
%f this is the case, then all of the genomes may have been subsequently diploidized, yielding the widespread and frequent chromosome numbers in this and closely related families, n=11--12, all presently considered to be diploid \citep{robertson_2011}.

% Part II: evidence against WGD in/near family
On the other hand, there is little evidence for the occurrence of diploidization after the origin of Solanaceae itself.
% R- This is redundant now we said it above
%First, the Ks-inferred duplication \citep{tomato2012} likely pre-dates the origin of the family (49 My, HPD 46--53 My; \citealt{sarkinen_2013}). 
%Thus, the species in our analyses could be ancestrally polyploid, but would have diploidized before the root of the family.
Studies comparing map-based genome synteny within the family find no evidence for recent diploidization (extensively discussed in \citet{wu_2010a}).
Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species. % including all of those in the relatively cytogenetically conserved x=12 group, which includes tomato, potato, eggplant, pepper, and tobacco.
Specifically, chromosome numbers and genome comparisons within the family (esp.\ the x=12 clade containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.

% R- removed the  inversion discussion I think the point was made above about the map base genome is sufficient. We don't need to describe more specifics

%\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ by as few as six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations.
%Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
%
% Part IIb: evidence from SI [might need more explaining?]
Furthermore, whole genome duplication in a eudicot lineage ancestral to Solanaceae would clash with the evidence that the homologous mechanism of SI, which has been present continually in many families \citep{igic_2006}, breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002, mcclure_2009}.
Most problematically in this context, it is unclear how to explain the maintenance of trans-generic polymorphism at the orthologous S-loci in Solanaceae and other families if SI was previously broken down by polyploidization.
%
Regardless of whether genome polyploidization, followed by widespread diploidization, is a dominant mechanism of genome evolution in Solanaceae, it is clear that more work is needed for a complete understanding of the joint evolution of ploidy and breeding systems.

### 2019-07-17 manuscript/introduction.tex
Fixed all todos and revised discussion issues

Species accumulate across the tree of life at different rates. 
One possible explanation for this phenomenon is that various traits differentially affect rates of diversification. 
Dramatic increases in phylogenetic and phenotypic data along with methodological advances have greatly accelerated the search for traits that influence diversification. 
Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a challenge \citep[\eg][]{maddison_2015, rabosky_2015, beaulieu_2016, rabosky_2017}. 
One difficulty is that speciation and extinction likely do not depend on a single character, so the biological and environmental contexts in which traits occur can lead to complex interactions that affect lineage diversification \citep{beaulieu_2016, caetano_2018, herrera_2018}.
Consequently, examining the association of only one character with diversification patterns can be misleading. 
We embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors to change diversification.
% E: We need a "Here, we..." here.  Otherwise, we've just told the reader "not one trait!" and then launch into one trait.
% Could also state that we newly test whether adding one more trait is worthwhile.
%B: I think you are correct, but it introduces a new concern: we now say N=1 is too few, so not why try N>2 characters? Why is N=2 the Goldilocks number? I didn't change anything but it's clear fodder for Discussion.

@@

In the present study, we employ an extended framework of state speciation and extinction models, which simultaneously model transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep[`SSE' models;][]{maddison_2007, fitzjohn_2012}.
% All the models proposed for this study are coded as graphical models and Bayesian inference was performed over the proposed models using RevBayes software \citep{hoehna_2016}.  % E: The models aren't drawn graphically in any of the figures (not in the revbayes sense, at least).  And we don't need methods in the intro.
%Although potentially very powerful, this approach can be prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
We start by fitting binary state speciation and extinction models to each ploidy and breeding system independently \citep{maddison_2007}.
We follow by fitting models that incorporate hidden states, which reduce the chance that an effect of one trait on diversification would be mistaken for the effect of an unobserved factor \citep{beaulieu_2016}.
We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually informative of the diversification process.
Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character.
Using the ploidy and breeding system model without hidden states, we quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
We also lump together states within these joint models of ploidy and breeding system in order to test whether increasing complexity from one trait to two traits significantly improves our understanding of the diversification process.
%B: I altered this line, but it is still a bit redundant given previous sentence.
%Our analyses evaluate combinations of macroevolutionary models to parse out the associations of ploidy and breeding system on speciation and extinction rates in Solanaceae, while simultaneously accounting for other processes that influence diversification.
%First, we present models in which the effects of ploidy and breeding system are considered separately and compare the inferences to previously published results. 
%Second, we add a hidden character to each model and investigate if state-dependent diversification in focal character retains its explanatory power, or whether the added hidden character obscures it.
Furthermore, we extend all the models involving ploidy to investigate the potential effects of including diploidization.
% And finally, we explore models of diversification using asymmetric rates between and within hidden states to investigate the sensitivity of the statistical inference to model assumptions.
Our results highlight the importance of considering non-additive effects of traits on net diversification rates  under the presence of unobserved factors, in order to detect strong biologically-driven processes dictating the diversification patterns.

>>>>

Species accumulate across the tree of life at different rates. 
One possible explanation for this phenomenon is that various traits differentially affect rates of diversification. 
Dramatic increases in phylogenetic and phenotypic data along with methodological advances have greatly accelerated the search for traits that influence diversification. 
Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a challenge \citep[\eg][]{maddison_2015, rabosky_2015, beaulieu_2016, rabosky_2017}. 
One difficulty is that speciation and extinction likely do not depend on a single character, so the biological and environmental contexts in which traits occur can lead to complex interactions that affect lineage diversification \citep{beaulieu_2016, caetano_2018, herrera_2018}.
Consequently, examining the association of only one character with diversification patterns can be misleading. 
Here, we embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors to change diversification. We also test whether adding one more trait and increasing model complexity is worthwhile.
% E: We need a "Here, we..." here.  Otherwise, we've just told the reader "not one trait!" and then launch into one trait.
% Could also state that we newly test whether adding one more trait is worthwhile.
%B: I think you are correct, but it introduces a new concern: we now say N=1 is too few, so not why try N>2 characters? Why is N=2 the Goldilocks number? I didn't change anything but it's clear fodder for Discussion.
% R. added Emma's suggetion

@@

In the present study, we employ an extended framework of state speciation and extinction models, which simultaneously model transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep[`SSE' models;][]{maddison_2007, fitzjohn_2012}.
% All the models proposed for this study are coded as graphical models and Bayesian inference was performed over the proposed models using RevBayes software \citep{hoehna_2016}.  % E: The models aren't drawn graphically in any of the figures (not in the revbayes sense, at least).  And we don't need methods in the intro.
% R- The graphical model coding is not about the figures but about how you code the nodes in the code. But I agree that it might be confusing especiallly for most botanists.
%Although potentially very powerful, this approach can be prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
We start by fitting binary state speciation and extinction models to each ploidy and breeding system independently \citep{maddison_2007}.
We follow by fitting models that incorporate hidden states, which reduce the chance that an effect of one trait on diversification would be mistaken for the effect of an unobserved factor \citep{beaulieu_2016}.
We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually informative of the diversification process.
Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character.
Using the ploidy and breeding system model without hidden states, we quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
We also aggregate  states within these joint models of ploidy and breeding system in order to test whether increasing complexity from one trait to two traits significantly improves our understanding of the diversification process.
%B: I altered this line, but it is still a bit redundant given previous sentence.R-fixed
%Our analyses evaluate combinations of macroevolutionary models to parse out the associations of ploidy and breeding system on speciation and extinction rates in Solanaceae, while simultaneously accounting for other processes that influence diversification.
%First, we present models in which the effects of ploidy and breeding system are considered separately and compare the inferences to previously published results. 
%Second, we add a hidden character to each model and investigate if state-dependent diversification in focal character retains its explanatory power, or whether the added hidden character obscures it.
Furthermore, we extend all the models involving ploidy to investigate the potential effects of including diploidization.
% And finally, we explore models of diversification using asymmetric rates between and within hidden states to investigate the sensitivity of the statistical inference to model assumptions.
Our results highlight the importance of considering non-additive effects of traits on net diversification rates  under the presence of unobserved factors, in order to detect strong biologically-driven processes dictating the diversification patterns.
